---
title: "MATH497-Best Draw?"
author: "Team Pbh"
date: "2023-03-30"
output: pdf_document
---

```{r}

rm(list=ls())
df1 <- read.csv('C:/Users/krtfe/Downloads/497CleanedSequentialData.csv')[-1]
df <- df1
df1 <- df1[-19]


df <- df[1:12]
df['Mean'] <- rowMeans(df[3:12])
df1['Mean'] <- rowMeans(df[3:12])
# KL divergence of pdf
dfRMNA <- na.omit(df)

dfRMNA <- dfRMNA[duplicated(dfRMNA$Address) == F,]

```


Quantile plots using qqplot() (comparing the quantiles of the plots vs. the quantiles of the mean):

```{r, message = FALSE}
library(EnvStats)
library(ggpubr)

# empty dataframe
dfQuant <- data.frame(Mean = rep(NA, nrow(dfRMNA)))

# 90th percentile for the mean
perc90 <- quantile(dfRMNA$Mean, probs = seq(0, 1, 0.9), na.rm = TRUE)[2]

# plotting the quantile data against each other
for (i in 3:12) {
  plotQuan <- qqplot(dfRMNA[,i], dfRMNA$Mean, plot = FALSE)
  dfQuant[paste('Draw_', i-2, sep = '')] <- plotQuan$x
  
  #plot
  print(ggplot() + 
          geom_point(aes(x=plotQuan$x, y=plotQuan$y)) + 
          geom_hline(yintercept = perc90, linetype = 'dashed', 
                     color = 'red') + 
          geom_abline(intercept = 0, slope = 1, color = 'blue') +
          ggtitle(paste('Quantiles of: Mean versus Draw', i-2)) +
          ylab('Quantiles of Mean') + xlab(paste('Quantiles of Draw',
                                                 i-2)) + 
          annotate(geom = 'text', 
                   label = '90th percentile line for mean', 
                   color = 'red', x = 137.5, y = perc90*1.5) +
          xlim(c(0, 175)))
}

# just getting the mean values
dfQuant$Mean <- plotQuan$y

```


Comparison of the quantiles of the mean versus the quantiles of the draws. 

Finding the R^2 values to determine the best relationship for the quantiles of the draws vs. the quantiles of the mean:

```{r}

for (i in 3:12) {
  b <- qqplot(df[,i], df$Mean, plot = FALSE)
  x <- b$x
  y <- b$y
  
  lg <- lm(y~x)
  summ <- summary(lg)$r.squared
  cat(paste('R^2 for draw ', i-2, ' vs. the mean: ', round(summ, 5), '\n', sep = ''))
}

```


From this, the 8th draw gives the highest R^2 for an R^2 value of 0.97874


Boxplots of the mean versus the draws, just general visualizations:

```{r, message = FALSE}

library(ggplot2)
library(reshape2)

restdf <- dfRMNA[3:13]
colnames(restdf) <- c('1', '2', '3', '4', '5', 
                      '6', '7', '8', '9', '10', 'Mean') 
a <- melt(restdf, na.rm = TRUE)
aplot <- ggplot(a[a$value < 40,], 
       aes(x = variable, y = value, fill = variable),
       ) + geom_boxplot()

aplot + 
  scale_fill_brewer(palette="Set3") + ylab('Lead ppb Values') + 
  ggtitle('Boxplots of the 10 Individual Draws and the Mean') + 
  scale_x_discrete(name = 'Draws and Mean', 
                   limits = c('1', '2', '3', '4', '5', 'Mean', '6', 
                              '7', '8', '9', '10') ) + 
  geom_hline(yintercept = 15, linetype = 'dashed', color = 'red') + 
  annotate(geom = 'text', 
           label = 'Lead Action Level - 15 ppb', size = 4.2,
           color = 'red', x = 3, y = 16.5) + 
  theme_bw()

```


KL Divergence, Sum of Absolute Value, and SSE of the CDFs (plots and dataframe)

```{r, message = FALSE, warning = FALSE}

# pdf of the mean
pM <- hist(dfRMNA$Mean, breaks = seq(0, max(dfRMNA$Mean), length.out = 100), plot = FALSE)$counts / nrow(dfRMNA)

# cdf of the mean
cdfM <- cumsum(pM)

# empty lists for computation
KLPlist <- c()
absvalL <- c()
ssevalL <- c()

# computing the differences for each type and each draw
for (i in 3:12) {
  # pdf of the draw
  p1 <- hist(unlist(dfRMNA[i]), 
                    breaks = seq(0, max(dfRMNA[i]),
                                 length.out=100), 
             plot = FALSE)$counts / nrow(dfRMNA)
  
  # cdf of the draw
  cdf1 <- cumsum(p1)
  
  # differences
  absval <- sum(abs(cdfM-cdf1))
  sseval <- sum((abs(cdfM-cdf1))^2)
  KLval <- max(abs(cdfM-cdf1))
  
  # add to list
  KLPlist <- c(KLPlist, KLval)
  absvalL <- c(absvalL, absval)
  ssevalL <- c(ssevalL, sseval)
} 

# plot of the SSE for the draws vs the mean
ggplot() + 
  geom_point(aes(x=1:10, y=ssevalL)) + 
  geom_line(aes(x=1:10, y=ssevalL)) + 
  ggtitle('Sum of Squared Errors for Mean CDF and Draw # CDF') + 
  ylab('SSE Value') + 
  ylim(c(0, 2)) + 
  scale_x_discrete(name = 'Draw #', limits = c(1:10)) +
  geom_vline(xintercept = 8, size = 1,
             linetype = 'dashed', color = 'maroon3') +
  geom_vline(xintercept = 5, size = 1, 
             linetype = 'dashed', color = 'darkorchid3') +
  geom_vline(xintercept = 1, size = 1,
             linetype = 'dashed', color = 'dodgerblue2') + 
  annotate(geom = 'text', 
           label = 'Minimum SSE Value', angle = 90,
           color = 'maroon3', x = 8*0.95, y = 1.1) +
  annotate(geom = 'text', 
           label = 'Proposed LCR Draw', angle = 90,
           color = 'darkorchid3', x = 4.7, y = 1.1) +
  annotate(geom = 'text', 
           label = 'Current LCR Draw', angle = 90,
           color = 'dodgerblue3', x = 0.7, y = 1.1) +
  theme_bw()

# plot of the KL-divergence for the draws vs. the mean
ggplot() + 
  geom_point(aes(x=1:10, y=KLPlist)) + 
  geom_line(aes(x=1:10, y=KLPlist)) + 
  ggtitle('Kullback-Leibler for Mean and Draw (cdfs)') + 
  ylab('KL Value') + 
  ylim(c(0, 0.7)) + 
  scale_x_discrete(name = 'Draw #', limits = c(1:10)) +
  geom_vline(xintercept = 8, size = 1,
             linetype = 'dashed', color = 'maroon3') +
  geom_vline(xintercept = 5, size = 1, 
             linetype = 'dashed', color = 'darkorchid3') +
  geom_vline(xintercept = 1, size = 1,
             linetype = 'dashed', color = 'dodgerblue2') + 
  annotate(geom = 'text', 
           label = 'Minimum KL Value', angle = 90,
           color = 'maroon3', x = 8*0.95, y = 0.35) +
  annotate(geom = 'text', 
           label = 'Proposed LCR Draw', angle = 90,
           color = 'darkorchid3', x = 4.7, y = 0.35) +
  annotate(geom = 'text', 
           label = 'Current LCR Draw', angle = 90,
           color = 'dodgerblue3', x = 0.7, y = 0.35) +
  theme_bw()

# plot for the sum of absolute error for the draws vs the mean
ggplot() + 
  geom_point(aes(x=1:10, y=absvalL)) + 
  geom_line(aes(x=1:10, y=absvalL)) + 
  ggtitle('Sum of Absolute Value for Mean and Draw (cdfs)') + 
  ylab('Sum of Absolute Value') + 
  ylim(c(0, 7.5)) + 
  scale_x_discrete(name = 'Draw #', limits = c(1:10)) +
  geom_vline(xintercept = 8, size = 1,
             linetype = 'dashed', color = 'maroon3') +
  geom_vline(xintercept = 5, size = 1, 
             linetype = 'dashed', color = 'darkorchid3') +
  geom_vline(xintercept = 1, size = 1,
             linetype = 'dashed', color = 'dodgerblue2') + 
  annotate(geom = 'text', 
           label = 'Minimum Sum of Absolute Value', angle = 90,
           color = 'maroon3', x = 8*0.95, y = 4) +
  annotate(geom = 'text', 
           label = 'Proposed LCR Draw', angle = 90,
           color = 'darkorchid3', x = 4.7, y = 4) +
  annotate(geom = 'text', 
           label = 'Current LCR Draw', angle = 90,
           color = 'dodgerblue3', x = 0.7, y = 4) +
  theme_bw()

# data frame for direct comparison
lnormVal <- data.frame('Absolute Error' = absvalL,
                       'Sum of Squared Errors' = ssevalL,
                       'Kullback Leibler' = KLPlist)
print(lnormVal)

```


Here, the 8th draw's distribution gives the lowest KL number. Ideally, the closer the number is to 0, the more similar the two distributions are to each other. This suggests that the 8th draw best matches the mean distribution.


The quantiles, all put together in one singular plot:

```{r, message = FALSE}

dfM <- melt(dfQuant, id.vars = 'Mean')
ggplot(dfM, aes(x = value, y = Mean, colour = variable)) + 
  geom_point() + 
  xlim(c(0, 100)) + 
  ylim(c(0, 100)) + 
  geom_hline(yintercept = perc90, linetype = 'dashed', color = 'red') + 
  annotate(geom = 'text', 
           label = '90th percentile line for mean', 
           color = 'red', x = 80, y = perc90*1.1825) + 
  geom_abline(intercept = 0, slope = 1) +
  ggtitle(paste('Quantiles of: Mean versus the Draws')) + 
  ylab('Quantiles of Mean') + 
  xlab('Quantiles of the Draws') +
  theme_bw()


```

Visualization of the PDFs

```{r, message = FALSE, warning = FALSE}
library(sm)

meltedDens <- melt(dfRMNA[c('X1st.Draw', 'X5th.Draw', 'Mean', 'X8th.Draw')])
ggplot(meltedDens, aes(x = value, fill = variable)) + 
  geom_density(alpha = 0.5, adjust = 1/5) +
  xlim(c(0, 50)) +
  xlab('ppb value') +
  scale_fill_brewer('Variable', palette = 'PiYG',
                    labels = c('1st Draw', '5th Draw', 'Mean', '8th Draw')) +
  theme_bw() +
  ggtitle('PDFs of the Mean, the 1st Draw, the 8th Draw, and the 5th Draw')

  
```


Visualization of the CDFs

```{r, message = FALSE, warning = FALSE}

p1 <- hist(dfRMNA$X1st.Draw, 
           breaks = seq(0, max(dfRMNA$X1st.Draw),
                               length.out=100), 
           plot = FALSE)$counts / nrow(dfRMNA)
cdf1 <- cumsum(p1)

p5 <- hist(dfRMNA$X5th.Draw, 
                  breaks = seq(0, max(dfRMNA$X5th.Draw),
                               length.out=100), 
             plot = FALSE)$counts / nrow(dfRMNA)
cdf5 <- cumsum(p5)

p8 <- hist(dfRMNA$X8th.Draw, 
           breaks = seq(0, max(dfRMNA$X8th.Draw),
                               length.out=100), 
           plot = FALSE)$counts / nrow(dfRMNA)
cdf8 <- cumsum(p8)

meltedDens <- melt(data.frame(cdf1, cdf5, cdf8, cdfM, 'x' = c(1:99)), 
                   id.vars = 'x')
ggplot(meltedDens, aes(y = value, x = x, color = variable)) + 
  geom_line() + 
  ggtitle('CDFs of the Mean, the 1st Draw, the 8th Draw, and the 5th Draw') +
  # geom_area(aes(fill = variable, group = variable),
  #           alpha = 0.25, position = 'identity') +
  scale_color_hue(name="Variables",
                  labels=c('1st Draw', '5th Draw', '8th Draw', 'Mean', 
                           '6th Draw')) +
  xlim(c(0, 50)) + 
  theme_bw()

```




